Bayesian Exploration in Deep Reinforcement Learning

### 2022-02-17 papers/main.tex
TEMPORARY DEPLOY COMMIT

%% Some suggested packages, as needed:
\usepackage{natbib} % has a nice set of citation styles and commands
    \bibliographystyle{plainnat}
    \renewcommand{\bibsection}{\subsubsection*{References}}
\usepackage{mathtools} % amsmath with fixes and additions
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
\usepackage{multicol}
\usepackage{tikz} % nice language for creating drawings and diagrams
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{bm}
\usepackage{subcaption}

@@

% \setlength{\tabcolsep}{12pt}

@@

\section{Introduction}
A reinforcement learning environement is modelled as a Markov decision process (MDP)
\(M = \langle \state, \action, r, P, P_0, \gamma \rangle\), where \(\state\) is the
state space and \(\action\) is the set of available actions. At time \(t=0\) a state
\(s_0\) is sampled from the distribution \(P_0(\cdot)\). At each timestep an action
\(a_t \sim \pi(\cdot \vert s_t)\) is selected and the agent transitions to a new state
state \(s_{t+1} \sim P(\cdot \vert s_t, a_t)\). A scalar reward
\(r_t = r(\cdot \vert s_t, a_t)\) is observed. As the agent and environement
interact in a sequence of time steps, a history of observations
\(\mathcal{H}_t = (s_0, a_0, r_0, s_1, a_1, r_1, \dots, s_t, a_t, r_t)\) is collected.
The goal is to find a policy \(\pi^\star\), such that sampling actions
\(a \sim \pi^*(\cdot \vert s)\) maximizes the expected accumulated and discounted future reward,
\(J^\pi \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \right]\). An efficient
agent must be able to learn from the data it collects, but since the data is
dependent on the policy, it must also prioritize to explore states and actions that
the agent can learn a lot from.

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward of following an optimal policy.
A learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy, is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is Thompson sampling~\citep{thompson_likelihood_1933}.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution and selecting the action that gave the highest
reward. Bayesian methods have also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

The \(Q\)-function, \(Q^\pi(s,a)\), is defined as the expected reward of taking action
\(a\) in state \(s\) and then following policy \(\pi\) thereafter:
\(Q^\pi(s_0, a_0) \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \vert s_0=S, a_0=A \right]\).
The Bellman operator on \(Q^\pi\) is defined as
\(\B[Q^\pi](s_t,a_t) \defeq \E_{P(s_{t+1} \vert s_t, a_t)\pi(a_{t+1} \vert s_t)}
\left[ r(s_t, a) + \gamma Q^\pi(s_{t+1}, a_{t+1}) \right]\). With the
\(Q\)-function we can define a policy that always picks the action with the 
highest \(Q\)-value in any state. In deep RL, we model \(Q\) with a
neural network \(\hat{Q}_\omega\). One way to facilitate exploration in this
policy is to introduce uncertainty in the
\(Q\)-value function.

@@

Before we dwelve into the background, we need to presicise some notation. Most of the
theory will be based on stochastic functions. The stochastic process we are interested is
generating value functions on an MDP, and exists on the probability space on the probability
space \((\Omega, \mathcal{F}, P)\). It can be written as
\(\{Q(\bm{s}, \bm{a}) : (\bm{s}, \bm{a}) \in \state \times \action \}\). For
any \(\omega \in \Omega\), \(Q(\cdot, \cdot, \omega)\) is a sample function
mapping \(\state \times \action \rightarrow \mathbb{R}\). To simplify notation
we will denote sample functions as \(f\). For a set of multiple sampled functions we will
use \(\f\), where \(\f(\X)\) means each sampled function evaluated at the same point,
i.e. \(\{f(\X) \mid f \in \f\}\).

@@

\subsection{Details and Hyperparameters}
We compare our method with NoisyNet~\citep{fortunato_noisy_2019} on four different
environments in the OpenAI Gym~\citep{brockman_openai_2016-2}. The hyperparameters
are the same as those described in \cite{han_nrowan-dqn_2020}. Our method additionally
has hyperparameters related to the functional KL-divergence. These are reported in
Table~\ref{tab:hyperparams}.

@@

\begin{align}
    p(\bm{s},a) &\sim \mathcal{N}(\bm{\mu}, \sigma I ), \text{ where}\\
    \bm{\mu} &= \lambda_1 + \lambda_2 \mathrm{A}_a \bm{s} \nonumber
\end{align}

% \begin{table}
%     \centering
%     \caption{Prior parameters}\label{tab:hyperparams}
%     \begin{tabular}{cccl}
%       \toprule % from booktabs package
%       Environment & \(\lambda_1\) & \(\lambda_2\) & \(\mathrm{A}\) \\
%       \midrule % from booktabs package
%       Cartpole & 0 & 100 & \(\begin{bmatrix} 0 & 0 & 1 & 1 \\
%                                              0 & 0 & -1 & -1
%                              \end{bmatrix}\)\\
%       MountainCar & 100 & 100 & \(\begin{bmatrix} 0 & -1 \\
%                                                   0 & 0 \\
%                                                   0 & 1
%                                 \end{bmatrix}\)\\
%       Acrobot & 100 & 100 & \(\begin{bmatrix} 0 & -1 & 0 & -1 & 0 & 0 \\
%                                                0 & 0 & 0 & 0 & 0 & 0 \\
%                                                0 & 1 & 0 & 1 & 0 & 0
%                             \end{bmatrix}\)\\
%       \bottomrule % from booktabs package
%     \end{tabular}
% \end{table}

>>>>

%% Some suggested packages, as needed:
\usepackage{natbib} % has a nice set of citation styles and commands
    \bibliographystyle{plainnat}
    \renewcommand{\bibsection}{\subsubsection*{References}}
\usepackage{mathtools} % amsmath with fixes and additions
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
\usepackage{multicol}
\usepackage{tikz} % nice language for creating drawings and diagrams
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{bm}
\usepackage{subcaption}
\usepackage{etoolbox}
\usepackage{lmodern}
\usepackage{siunitx}

@@

\sisetup{detect-weight,mode=text}
% for avoiding siunitx using bold extended
\renewrobustcmd{\bfseries}{\fontseries{b}\selectfont}
\renewrobustcmd{\boldmath}{}
\newrobustcmd{\tB}{\bfseries}

\setlength{\tabcolsep}{2pt}

@@

\section{Introduction}
A reinforcement learning environement is modelled as a Markov decision process (MDP)
\(M = \langle \state, \action, r, P, P_0, \gamma \rangle\), where \(\state\) is the
state space and \(\action\) is the set of available actions. At time \(t=0\) a state
\(s_0\) is sampled from the distribution \(P_0(\cdot)\). At each timestep an action
\(a_t \sim \pi(\cdot \vert s_t)\) is selected and the agent transitions to a new state
state \(s_{t+1} \sim P(\cdot \vert s_t, a_t)\). A scalar reward
\(r_t = r(s_t, a_t)\) is observed. As the agent and environement
interact in a sequence of time steps, a history of observations
\(\mathcal{H}_t = (s_0, a_0, r_0, s_1, a_1, r_1, \dots, s_t, a_t, r_t)\) is collected.
The goal is to find a policy \(\pi^\star\), such that sampling actions
\(a \sim \pi^*(\cdot \vert s)\) maximizes the expected accumulated and discounted future reward,
\(J^\pi \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \right]\). An efficient
agent must be able to learn from the data it collects, but since the data is
dependent on the policy, it must also prioritize to explore states and actions that
the agent has little information on.

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward of following an optimal policy.
A learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy, is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is Thompson sampling~\citep{thompson_likelihood_1933}.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution and selecting the action that gave the highest
reward. Bayesian methods have also been shown to behave efficiently with respect
to cumulative regret \citep{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

The \(Q\)-function, \(Q^\pi(s,a)\), is defined as the expected reward of taking action
\(a\) in state \(s\) and then following policy \(\pi\) thereafter:
\(Q^\pi(s_0, a_0) \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \vert S=s, A_0=a \right]\).
The Bellman operator on \(Q^\pi\) is defined as
\(\B[Q^\pi](s_t,a_t) \defeq \E_{P(s_{t+1} \vert s_t, a_t)\pi(a_{t+1} \vert s_t)}
\left[ r(s_t, a) + \gamma Q^\pi(s_{t+1}, a_{t+1}) \right]\). With the
\(Q\)-function we can define a policy that always picks the action with the 
highest \(Q\)-value in any state. In deep RL, we model \(Q\) with a
neural network \(\hat{Q}_\omega\). One way to facilitate exploration in this
policy is to introduce uncertainty in the
\(Q\)-value function.

@@

Before we dwelve into the background, we need to define some notation. Most of the
theory will be based on stochastic functions. The stochastic process we are interested is
generating value functions on an MDP, and exists on the probability space on the probability
space \((\Omega, \mathcal{F}, P)\). It can be written as
\(\{Q(\bm{s}, \bm{a}) : (\bm{s}, \bm{a}) \in \state \times \action \}\). For
any \(\omega \in \Omega\), \(Q(\cdot, \cdot, \omega)\) is a sample function
mapping \(\state \times \action \rightarrow \mathbb{R}\). To simplify notation
we will denote sample functions as \(f\). For a set of multiple sampled functions we will
use \(\f\), where \(\f(\X)\) means each sampled function evaluated at the same point,
i.e. \(\{f(\X) \mid f \in \f\}\).

@@

\subsection{Details and Hyperparameters}
We compare our method with NoisyNet~\citep{fortunato_noisy_2019} on four different
environments in the OpenAI Gym~\citep{brockman_openai_2016-2}. The hyperparameters
are the same as those described  \cite{han_nrowan-dqn_2020}. Our method additionally
has hyperparameters related to the functional KL-divergence. These are reported in
Table~\ref{tab:hyperparams}.

@@

\begin{align}
    p(\bm{s},a) &\sim \mathcal{N}(\bm{\mu}, \sigma I ), \text{ where}\\
    \bm{\mu} &= \lambda_1 + \nu \lambda_2 \mathrm{A}_a \bm{s} \nonumber
\end{align}

\begin{table}
    \centering
    \caption{Prior parameters}\label{tab:prior_hyperparams}
    \begingroup
    \setlength\arraycolsep{3pt}
    \begin{tabular}{lccc}
      \toprule % from booktabs package
      Environment & \(\lambda_1\) & \(\lambda_2\) & \(\mathrm{A}\) \\
      \midrule % from booktabs package
      Cartpole & 0 & 100 & \(\begin{bmatrix} 0 & 0 & 1 & 1 \\
                                             0 & 0 & -1 & -1
                             \end{bmatrix}\)\\\addlinespace[0.3em]
      MountainCar & -100 & 100 & \(\begin{bmatrix} 0 & -1 \\
                                                  0 & 0 \\
                                                  0 & 1
                                \end{bmatrix}\)\\\addlinespace[0.3em]
      Acrobot & -100 & 100 & \(\begin{bmatrix} 0 & -1 & 0 & -1 & -0.2 & -0.1 \\
                                               0 & 0 & 0 & 0 & 0 & 0 \\
                                               0 & 1 & 0 & 1 & 0.2 & 0.1
                            \end{bmatrix}\)\\
      \bottomrule % from booktabs package
    \end{tabular}
    \endgroup
\end{table}

\begin{table}
    \centering
    \caption{Number of episodes to beat select environments.}\label{tab:prior_results}
    \begin{tabular}{cccc}
        \toprule
        Prior & Cartpole & MountainCar & LunarLander\\
        \midrule
        \(\nu = 1, \sigma = 1\)   & \tB 14  & 100 &\\
        \(\nu = 1, \sigma = 10\)  &     33  & 1700 &\\
        \(\nu = 1, \sigma = 50\)  &     113 & 1250 &\\
        Flat prior                &     117 & 1650 &\\
        \(\nu = -1, \sigma = 50\) &     177 & &\\
        \(\nu = -1, \sigma = 20\) &     232 & &\\
        \(\nu = -1, \sigma = 10\) &     --- & &\\
        \midrule
        NoisyNet                  &     219 & 1700 &\\
        \bottomrule
    \end{tabular}
\end{table}
